Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-10-04 psy_peds_3.Rmd
adding time series analysis

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
time_period <-"months" #possible values: days, weeks, months
```

@@

clinical_filtered$time_p <- NA
if( time_period == "weeks"){
  clinical_filtered$time_p <- clinical_filtered$weeks
}else if(time_period == "days"){
  clinical_filtered$time_p <- clinical_filtered$date
}else if(time_period == "months"){
  clinical_filtered$time_p <- clinical_filtered$month
  }
```

@@

>>>>

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
time_period <-"months" #possible values: days, weeks, months
```

@@

if( time_period == "weeks"){
  clinical_filtered$time_p <- clinical_filtered$weeks
}else if(time_period == "days"){
  clinical_filtered$time_p <- clinical_filtered$date
}else if(time_period == "months"){
  clinical_filtered$time_p <- clinical_filtered$month
}

@@

## Interrupted Time-Series Analysis

```{r}
# Model without seasonality
time_series_analysis_plotting <- function(model, df_count_psy, count_col, clearance_period = FALSE) {
  output_model <- broom::tidy(model)
  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
  pred_model <- predict(model, patient_count_monthly_psy, type = "response", interval = "confidence") %>% as.data.frame()
  pred_model_b <- predict(model, mutate(patient_count_monthly_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
    as.data.frame()
  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
  pred_model <- bind_cols(df_count_psy, pred_model, pred_model_b)
  if (clearance_period) {
    pred_model$clearance_period <- ifelse(pred_model$time_p < as.Date(cut(start_clear_period, breaks = time_period)) ,
                                          "before_pandemic",
                                          ifelse(pred_model$time_p > as.Date(cut(end_clear_period, breaks = time_period)),
                                                 "during_pandemic",
                                                 "during_clearance")
    )
    pred_model$lwr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$lwr)
    pred_model$upr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$upr)
    pred_model$fit_wo_intervention <- ifelse(pred_model$time_p >= max(pred_model$time_p[pred_model$clearance_period == "before pandemic"]),
                                             pred_model$fit_wo_intervention,
                                             NA)
  } else {
    pred_model$clearance_period <- pred_model$period
  }
  print(pred_model$clearance_period)
  plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
    geom_point(aes(colour = clearance_period)) +
    geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
    geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
    scale_color_manual(name = "Time Series projection", values = c("before_pandemic" = "black",
                                                                   "Expected mean w/o pandemic" = "red",
                                                                   "during_pandemic" = "black",
                                                                   "during_clearance" = "white")) +
    annotate("text", x = as.Date("2019-10-01"), y = max(pred_model[[count_col]]) + 5, label = annotation) +
    theme(legend.position = "bottom")

  if (clearance_period) {
    plot +
      geom_vline(xintercept = as.Date(max(pred_model$time_p[pred_model$clearance_period == "before_pandemic"])), linetype = "dashed") +
      geom_vline(xintercept = as.Date(min(pred_model$time_p[pred_model$clearance_period == "during_pandemic"])), linetype = "dashed")
  } else {
    plot + geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")
  }
}
```

```{r}
# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
patient_count_monthly_psy <- count_icd_psy %>%
  filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
           time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
  mutate(time = seq_along(time_p),
         month_year = format(time_p, "%m")
  )
start_clear_period <- as.Date("2020-03-15")
end_clear_period <- as.Date("2020-05-15")
patient_count_monthly_psy_clear <- patient_count_monthly_psy %>%
  mutate( count_psy = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                             count_psy,
                             NA)
  )
```

```{r}
ts_linear_count <- lm(count_psy ~ period * time, patient_count_monthly_psy)
ts_linear_count_clear <- lm(count_psy ~ period * time, patient_count_monthly_psy_clear)
ts_seasonal_count <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy)
ts_seasonal_count_clear <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy_clear)
```

```{r}
time_series_analysis_plotting(ts_linear_count,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

```{r}
time_series_analysis_plotting(ts_linear_count_clear,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

```{r}
time_series_analysis_plotting(ts_seasonal_count,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```

```{r}
time_series_analysis_plotting(ts_seasonal_count_clear,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```
